---
title: "Group contributions to OSM"
format: 
  dashboard:
    # orientation: columns
    orientation: rows
    # scrolling: true
    self-contained: true
    nav-buttons:
        - icon: github
          href: https://github.com/WarwickCIM/OSMdashboard
          
params:
  use_db_overlay: false #this allows the database to be used to populate the .csv files
  db_path: "D:/GitHub/OSMdashboard/database/osm_changesets.duckdb" #please change if its elsewhere
  hashtags: ""  #formatted as "#HOT, #MissingMaps"
  usernames: ""# formatted as "alice123,bob_mapper"
  start: "" # formatted as "2023-01-01"
  end: ""   # formatted as "2023-01-01"
  bbox: "" #formatted a: "min_lat,min_lon,max_lat,max_lon" e.g "51.2,-0.5,51.8,0.3"     
---

```{r user input}
# Uncomment to update datasets.
# source(file = "data_retrieval.R")

# Add folder name if dashboard is not in the root of the project. Add trailing /
base_path <- ""

```


```{r setup}
#| echo: false
#| warning: false
#| output: false

library(dplyr)
library(forcats)
library(ggplot2)
library(here)
library(knitr)
library(leaflet)
library(leaflet.extras)
library(lubridate)
library(OSMdashboard)
library(plotly)
library(readr)
library(sf)
library(stringr)
library(tidyr)
library(treemapify)
library(DT)

# GGplot configs --------------------------------------------------------------

# theme_set(
#   theme_minimal(
#     base_size = 14) +
#   theme(
#     plot.title = element_text(face = "bold"),
#     plot.title.position = "plot",
#     legend.position = "top"
#   )
# )

# Heatmap theme for Changeset and WikiEdit distribution
theme_heatmap <- theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(
      family = "Arial",
      face = "bold",
      size = 16,
      hjust = 0.5
    ),
    axis.text.x = element_text(family = "Arial", size = 8,
                               angle = 45, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(family = "Arial", size = 8),
    legend.position = "none"
  )

theme_user_stats <- theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.line = element_line(colour = 'grey20', size = 0.5),
        legend.position = "none",
        plot.margin = margin(5,20,5,20))

## Colour Scheme: https://colorbrewer2.org/#type=diverging&scheme=PuOr&n=4

# Colours for marks
colour_primary_mark <- "#e66101"
colour_secondary_mark <- "#5e3c99"
colour_tertiary_mark <- "#2d004b"
colour_fill <- "#fdb863"

```


```{python}
#| label: db-overlay
#| echo: false
#| cache: false
#| message: false 
#| warning: false

import os, sys, subprocess

# make sure Python can see the python folder in the main directory
sys.path.insert(0, os.path.join(os.getcwd(), "../python"))

#ensure all python libraries are installed
def ensure(pkg):
    try: __import__(pkg)
    except ImportError: subprocess.check_call([sys.executable, "-m", "pip", "install", pkg])

for p in ("duckdb","pandas","pyarrow"): ensure(p)

#import python libraries
import pandas as pd
import duckdb

#import helper module after installing the python libraries
import overlay

#import helper function from module
from overlay import build_overlay

p = params if 'params' in globals() else {}

db_path_raw = p.get("db_path") or os.getenv("DUCKDB_PATH", "../database/osm_changesets.duckdb")

db_path = db_path_raw if os.path.isabs(db_path_raw) else os.path.abspath(os.path.join(os.getcwd(), db_path_raw))

# read params with sane fallbacks
usernames = [u.strip() for u in (p.get("usernames") or "").split(",") if u.strip()] or None
hashtags  = [h.strip() for h in (p.get("hashtags")  or "").split(",") if h.strip()] or None
start     = p.get("start") or None
end       = p.get("end") or None
bbox      = tuple(map(float, p.get("bbox").split(","))) if p.get("bbox") else None

#build the overlay with information from the database
build_overlay(
    db_path=db_path,
    queries_path="../python/queries.py",
    group_users_csv="data/metadata/group_users.csv",
    group_info_csv="data/metadata/group_info.csv",
    out_dir="data/db_overlay",
    usernames=usernames,
    hashtags=hashtags,
    start=start,
    end=end,
    bbox=bbox,
)
```


```{r}
#| label: setup-data
#| echo: false

# Detect base path automatically (works from root or from subfolders like database_test/)
base_path <- if (dir.exists("data/metadata")) "" else
             if (dir.exists("../data/metadata")) "../" else ""

use_db <- isTRUE(params$use_db_overlay)

data_root_candidates <- c("data", "../data")
data_root <- NULL
for (d in data_root_candidates) {
  if (file.exists(file.path(d, "metadata", "group_info.csv"))) {
    data_root <- d
    break
  }
}
if (is.null(data_root)) {
  stop("group_info.csv not found in ", paste(data_root_candidates, collapse = " or "),
       "\nwd: ", getwd())
}

# helper that returns NULL if file doesn't exist
read_csv_if <- function(path) {
  if (file.exists(path)) readr::read_csv(path, show_col_types = FALSE) else NULL
}

group_info <- readr::read_csv(file.path(data_root, "metadata", "group_info.csv"),
                              show_col_types = FALSE)

#raed either from overlay or raw depending on use_db
if (use_db) {
  cs_path   <- file.path(data_root, "db_overlay", "changesets_subset.csv")
  cst_path  <- file.path(data_root, "db_overlay", "changesets_tags_subset.csv")
  cons_path <- file.path(data_root, "db_overlay", "contributions_summary_subset.csv")
} else {
  cs_path   <- file.path(data_root, "raw", "changesets.csv")
  cst_path  <- file.path(data_root, "raw", "changesets_tags.csv")
  cons_path <- file.path(data_root, "raw", "contributions_summary.csv")
}

#load what we can
contributions_summary <- read_csv_if(cons_path)
changesets            <- read_csv_if(cs_path)
changesets_tags       <- read_csv_if(cst_path)

# wiki is not produced by the overlay - keep loading from RAW via data_root
wiki_contributions <- read_csv_if(file.path(data_root, "raw", "wiki_contributions.csv"))

if (is.null(changesets)) stop("Could not load changesets from: ", cs_path)

if ("created_at" %in% names(changesets)) {
  changesets$created_at <- lubridate::ymd_hms(changesets$created_at, quiet = TRUE)
}

#if contributions_summary is missing, create a minimal shell so later code works
if (is.null(contributions_summary)) {
  u <- sort(unique(changesets$user))
  contributions_summary <- tibble::tibble(
    user = u,
    map_changesets = 0, map_notes = 0, diary = 0, comments = 0, traces = 0, wiki_edits = 0,
    account_age = 0
  )
}

if (use_db) {
  # Ensure lon/lat columns exist for centroids
  if (!all(c("lon","lat") %in% names(changesets))) {
    changesets <- dplyr::mutate(
      changesets,
      lon = (min_lon + max_lon) / 2,
      lat = (min_lat + max_lat) / 2
    )
  }

  # Clean and validate the bounding box values
  bbox_cols <- c("min_lat","min_lon","max_lat","max_lon")
  changesets <- changesets |>
    dplyr::mutate(
      dplyr::across(all_of(bbox_cols), ~ suppressWarnings(as.numeric(.))),
      ymin = pmin(min_lat, max_lat),
      ymax = pmax(min_lat, max_lat),
      xmin = pmin(min_lon, max_lon),
      xmax = pmax(min_lon, max_lon)
    )

  valid <- with(changesets,
    is.finite(xmin) & is.finite(ymin) & is.finite(xmax) & is.finite(ymax) &
    xmax > xmin & ymax > ymin &
    dplyr::between(xmin, -180, 180) & dplyr::between(xmax, -180, 180) &
    dplyr::between(ymin,  -90,  90) & dplyr::between(ymax,  -90,  90)
  )

  # Build polygons only for valid rows
  mk_one <- function(i) {
    xmn <- changesets$xmin[i]; ymn <- changesets$ymin[i]
    xmx <- changesets$xmax[i]; ymx <- changesets$ymax[i]
    sf::st_polygon(list(matrix(
      c(xmn, ymn,
        xmx, ymn,
        xmx, ymx,
        xmn, ymx,
        xmn, ymn),
      ncol = 2, byrow = TRUE
    )))
  }

  idx <- which(valid)
  if (length(idx)) {
    polys <- lapply(idx, mk_one)
    changesets_sf <- sf::st_as_sf(sf::st_sfc(polys, crs = 4326))
    # project to meters for area and write back per-row
    changesets$area_meters <- NA_real_
    changesets$area_meters[idx] <-
      as.numeric(sf::st_area(sf::st_transform(changesets_sf, 3857)))
  } else {
    changesets_sf <- sf::st_sf(geometry = sf::st_sfc(), crs = 4326)
    changesets$area_meters <- NA_real_
  }

  #centroids from lon/lat (already computed)
  changesets_sf_centroids <- changesets |>
    dplyr::filter(!is.na(lon), !is.na(lat)) |>
    sf::st_as_sf(coords = c("lon","lat"), crs = 4326, remove = FALSE)
} else {
  # RAW path: read existing polygons from the GPKG
  raw_gpkg <- file.path(data_root, "raw", "changesets.gpkg")
  changesets_sf <- sf::read_sf(raw_gpkg)
  changesets_sf_repaired <- sf::st_make_valid(changesets_sf)

  #centroids robust to any geometry
  centroids <- sf::st_centroid(changesets_sf_repaired)
  coords <- sf::st_coordinates(centroids)
  changesets_sf_centroids <- centroids
  changesets_sf_centroids$lon <- coords[, 1]
  changesets_sf_centroids$lat <- coords[, 2]

  # Area from repaired polygons
  changesets$area_meters <- as.numeric(
    sf::st_area(sf::st_transform(changesets_sf_repaired, 3857))
  )
}

#common derived scalars used later
n_users <- dplyr::n_distinct(changesets$user)
date_start <- as.Date(min(changesets$created_at, na.rm = TRUE))
date_end   <- as.Date(max(changesets$created_at, na.rm = TRUE))
n_days <- as.numeric(difftime(date_end, date_start, units = "days"))
contributions_n_changesets <- nrow(changesets)
```

```{r}
###Back-fill from precomputed file, then compute contributions_type
#| label: api-backfill-profile-counts
#| echo: false
#| message: true
#| warning: false

# We back-fill diary/notes/traces/comments from data/raw/osm_user_details.csv
# (written by data_retrieval.R)

must_cols <- c("map_notes","diary","comments","traces")
have_cols <- must_cols %in% names(contributions_summary)
missing_any <- any(!have_cols)

all_zero <- FALSE
if (!missing_any) {
  all_zero <- all(vapply(
    must_cols,
    function(col) all(replace(contributions_summary[[col]], is.na(contributions_summary[[col]]), 0) == 0),
    logical(1)
  ))
}

need_fill <- isTRUE(use_db) && (missing_any || all_zero)

if (need_fill) {
  # read the precomputed user details
  counts_path <- file.path(data_root, "raw", "osm_user_details.csv")
  if (file.exists(counts_path)) {
    api_counts <- readr::read_csv(counts_path, show_col_types = FALSE) |>
      dplyr::mutate(user = tolower(user))
  } else {
    # keep rendering even if the file is missing
    api_counts <- tibble::tibble(user = character())
  }

  # ensure required cols exist & are numeric, then merge
  for (col in must_cols) if (!col %in% names(api_counts)) api_counts[[col]] <- 0
  api_counts <- api_counts |>
    dplyr::mutate(dplyr::across(all_of(must_cols), ~as.numeric(replace(., is.na(.), 0)))) |>
    dplyr::select(user, all_of(must_cols))

  contributions_summary <- contributions_summary |>
    dplyr::select(-dplyr::any_of(must_cols)) |>
    dplyr::left_join(api_counts, by = "user") |>
    dplyr::mutate(dplyr::across(all_of(must_cols), ~as.numeric(replace(., is.na(.), 0))))
}

# compute contributions_type
contributions_type <- contributions_summary |>
  dplyr::select(-dplyr::starts_with("date"), -dplyr::any_of("account_age")) |>
  tidyr::pivot_longer(-user) |>
  dplyr::count(name, wt = value) |>
  dplyr::filter(name %in% c("comments","diary","map_changesets","map_notes","traces","wiki_edits"))
```

# Overview

## Title and description 

::: {.card .flow}

## `r group_info$name[1]`

`r group_info$description[1]`

:::


## Value boxes

```{r}
#| content: valuebox
#| title: "# users"
#| icon: people
#| color: primary
nrow(contributions_summary)
```

```{r}
#| content: valuebox
#| title: "Mean account age (years)"
#| icon: hourglass-split
#| color: primary
round(mean(contributions_summary$account_age), digits = 2)
```


Hours spent contributing (estimate based on time x changesets x features + time * written words + time * delete words + time for discussion + time for ...)

## Group Summary (Group Information, Group Description)


```{r}

#| message: false
#| warning: false

edits_tbl <- if (exists("edits")) edits else if (exists("elements")) elements else NULL
tags_tbl  <- if (exists("tags_summary")) tags_summary else NULL

```

## Some summaries {height=85%}

### User Stats {.tabset}

```{r}
#| title: "Account Age"

hist_account_age <- contributions_summary |>
  ggplot(aes(x = account_age)) +
  geom_histogram(
    breaks = seq(0, 20, by = 2),
    color = colour_primary_mark,
    fill = colour_fill
  ) +
  geom_vline(
    xintercept = mean(contributions_summary$account_age), 
    color = colour_secondary_mark, 
    linetype = 2, 
    lwd = 0.7
  ) + 
  annotate(
    "text", 
    x = mean(contributions_summary$account_age), 
    y = 3, 
    label = paste("Mean:", round(mean(contributions_summary$account_age), 2)),
    color = colour_secondary_mark, 
    fontface = "bold"
  ) + 
  scale_x_continuous(breaks = seq(0, 20, by = 5)) +
  labs(y = "# users", x = "Account's age (in years)") +
  theme_user_stats


```

```{r}
#| title: "Map changesets"

hist_map_changesets <- contributions_summary |>
  ggplot(aes(x = map_changesets)) +
  geom_histogram(
    # breaks = seq(0, max(contributions_summary), by = 2),
    color = colour_primary_mark,
    fill = colour_fill
  ) +
  geom_vline(
    xintercept = mean(contributions_summary$map_changesets), 
    color = colour_secondary_mark, 
    linetype = 2, 
    lwd = 0.7
  ) + 
  annotate(
    "text", 
    x = mean(contributions_summary$map_changesets), 
    y = 3, 
    label = paste("Mean:", round(mean(contributions_summary$map_changesets), 2)),
    color = colour_secondary_mark, 
    fontface = "bold"
  ) + 
  # scale_x_continuous(breaks = seq(0, 20, by = 5)) +
  labs(y = "# of users", x = "Changesets") +
  theme_user_stats

```


```{r}
#| title: "Map changesets"

hist_map_notes <- contributions_summary |>
  ggplot(aes(x = map_notes)) +
  geom_histogram(
    # breaks = seq(0, max(contributions_summary), by = 2),
    color = colour_primary_mark,
    fill = colour_fill
  ) +
  geom_vline(
    xintercept = mean(contributions_summary$map_notes), 
    color = colour_secondary_mark, 
    linetype = 2, 
    lwd = 0.7
  ) + 
  annotate(
    "text", 
    x = mean(contributions_summary$map_notes), 
    y = 3, 
    label = paste("Mean:", round(mean(contributions_summary$map_notes), 2)),
    color = colour_secondary_mark, 
    fontface = "bold"
  ) + 
  # scale_x_continuous(breaks = seq(0, 20, by = 5)) +
  labs(y = "# of users", x = "Notes") +
  theme_user_stats

```

```{r}
#| title: "Comment Bigrams"

hist_comments <- contributions_summary |>
  ggplot(aes(x = comments)) +
  geom_histogram(
    # breaks = seq(0, max(contributions_summary), by = 2),
    color = colour_primary_mark,
    fill = colour_fill
  ) +
  geom_vline(
    xintercept = mean(contributions_summary$comments), 
    color = colour_secondary_mark, 
    linetype = 2, 
    lwd = 0.7
  ) + 
  annotate(
    "text", 
    x = mean(contributions_summary$comments), 
    y = 3, 
    label = paste("Mean:", round(mean(contributions_summary$comments), 2)),
    color = colour_secondary_mark, 
    fontface = "bold"
  ) + 
  # scale_x_continuous(breaks = seq(0, 20, by = 5)) +
  labs(
    subtitle = "Comments are discussions around a particular changeset",
    x = "Comments",
    y = "# of users"
  ) +
  theme_user_stats

```


```{r}
#| title: "Diaries"

hist_diary <- contributions_summary |>
  ggplot(aes(x = diary)) +
  geom_histogram(
    # breaks = seq(0, max(contributions_summary), by = 2),
    color = colour_primary_mark,
    fill = colour_fill
  ) +
  geom_vline(
    xintercept = mean(contributions_summary$diary), 
    color = colour_secondary_mark, 
    linetype = 2, 
    lwd = 0.7
  ) + 
  annotate(
    "text", 
    x = mean(contributions_summary$diary), 
    y = 3, 
    label = paste("Mean:", round(mean(contributions_summary$diary), 2)),
    color = colour_secondary_mark, 
    fontface = "bold"
  ) + 
  # scale_x_continuous(breaks = seq(0, 20, by = 5)) +
  labs(y = "# of users", x = "Diary entries") +
  theme_user_stats

```


```{r}
#| title: "Traces"

hist_traces <- contributions_summary |>
  ggplot(aes(x = traces)) +
  geom_histogram(
    # breaks = seq(0, max(contributions_summary), by = 2),
    color = colour_primary_mark,
    fill = colour_fill
  ) +
  geom_vline(
    xintercept = mean(contributions_summary$traces), 
    color = colour_secondary_mark, 
    linetype = 2, 
    lwd = 0.7
  ) + 
  annotate(
    "text", 
    x = mean(contributions_summary$traces), 
    y = 3, 
    label = paste("Mean:", round(mean(contributions_summary$traces), 2)),
    color = colour_secondary_mark, 
    fontface = "bold"
  ) + 
  # scale_x_continuous(breaks = seq(0, 20, by = 5)) +
  labs(y = "# of users", x = "Traces") +
  theme_user_stats

```

```{r}
#| title: "Contributions distribution by users"
#| subtitle: "This provides info about how homogeneous or heterogeneous this group is"
#| layout-ncol: 2


ggplotly(hist_account_age) 
ggplotly(hist_map_changesets)
ggplotly(hist_map_notes)
ggplotly(hist_comments)
ggplotly(hist_diary)
ggplotly(hist_traces)

```

```{r}
#| title: "Facets"

contributions_summary |> 
  select(user, account_age, map_changesets, wiki_edits, comments, diary, map_notes, traces) |> 
  pivot_longer(-user) |> 
  mutate(name = as.factor(name),
         name = fct_relevel(name, "account_age", "map_changesets", "wiki_edits", "comments", "diary", "map_notes", "traces")) |> 
  ggplot(aes(x = value, fill = name)) +
  geom_histogram(
    # breaks = seq(0, max(contributions_summary), by = 2),
    # color = colour_primary_mark,
    # fill = colour_fill
  ) +
  geom_vline(
    xintercept = mean(~value), 
    # color = colour_secondary_mark, 
    linetype = 2, 
    lwd = 0.7
  ) + 
  facet_wrap(~name, ncol = 2, scales = "free_x")+ 
  labs(y = "", x ="") +
  # theme_minimal() +
  theme(
    axis.title.y=element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    panel.grid= element_blank(),
    legend.position = "none"
  )


```


### Contributions distribution

```{r}
#| title: "Contributions per type"

contributions_type |> 
  plot_ly(labels = ~name, values = ~n) |> 
  add_pie(hole = 0.6)

```


# Changesets {icon="fa-utensils"}

## Valueboxes

```{r}
#| content: valuebox
#| title: "# users with changesets"
#| icon: map
#| color: primary
contributions_summary |> 
  filter(map_changesets > 0) |> 
  nrow()
```

```{r}
#| content: valuebox
#| title: "# changesets"
#| icon: pencil
#| color: secondary
contributions_n_changesets
```

```{r}
#| content: valuebox
#| title: "changesets/day"
#| icon: pencil-fill
round(nrow(changesets)/n_days, digits = 2)
```

<!-- Create a boxplot that shows on the x axis number of changesets, on the Y axis number of features per changeset, and the size of the dot can be the area of the changeset. Do this for every member of the group. -->


## Map of contributions { height="300"}

```{r}
#| title: Where do they contribute?
#| padding: 0px
leaflet(options = leafletOptions(maxZoom = 12)) |> 
  addTiles() |>   # Add default OpenStreetMap map tiles
  addProviderTiles(providers$CartoDB.Positron) |> 
  # addProviderTiles(providers$Esri.WorldStreetMap) |> 
  # addPolygons(data = changesets_sf) |> 
  # addCircles(data = changesets_sf_centroids)
  addHeatmap(data = filter(changesets_sf_centroids, !is.na(lon)), lng = ~lon, lat = ~lat,
           blur = 10, radius = 20, minOpacity = 0.5 ) 
```


```{r}
#| title: When do they contribute?

# Modified - OSM_Aug_RA
# changesets_day_time <- changesets |> 
#   select(created_at) |> 
#   mutate(date = created_at,
#          weekday = weekdays(date),
#          weekday = forcats::fct_relevel(
#            as.factor(weekday), 
#            "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"),
#          time = hms::as_hms(date),
#          hour = hour(time)) |> 
#   count(weekday, hour) |> 
#   # Converting to wide to generate NAs
#   pivot_wider(names_from = hour, values_from = n) |>
#   pivot_longer(-weekday, names_to = "hour", values_to = "n") |> 
#   mutate(hour = as.numeric(hour)) 
# # mutate(n = replace_na(n, 0))
# 
# heatmap <- ggplot(changesets_day_time, aes(hour, weekday, fill= n)) + 
#   geom_tile(show.legend = FALSE) +
#   labs(y = NULL) + 
#   scale_fill_distiller(palette = "Greens", direction = 1) +
#   # scale_x_continuous(position = "top")  + # Does not work with plotly!
#   theme_minimal() 
# # theme(panel.background = element_rect(fill = 'lightgrey'),
# #       # panel.grid.minor = element_line(color = 'white', size = 2)
# #       )
# 
# heatmap_interactive <- ggplotly(heatmap)
# 
# heatmap_interactive

changesets_day_time <- changesets |>
  select(created_at) |>
  mutate(
    date = created_at,
    month = factor(as.integer(format(date, "%m")), levels = 1:12, labels = month.abb, ordered = TRUE),
    weekday = weekdays(date),
    weekday = forcats::fct_relevel(
      as.factor(weekday),
      "Monday",
      "Tuesday",
      "Wednesday",
      "Thursday",
      "Friday",
      "Saturday",
      "Sunday"
    )
  ) |>   count(month, weekday)

changeset_heatmap <-
  ggplot(changesets_day_time, aes(month, weekday, fill = n)) +
  geom_tile() +
  labs(x = NULL, y = NULL) +
  scale_fill_gradient2(
    low = "#ffffff",
    mid = "#d8daeb",
    high = "#2d004b",
    name = "Number of Changes"
  ) +
  scale_y_discrete(limits = rev) +
  #coord_fixed() +
  theme_heatmap

changeset_heatmap_interactive <- ggplotly(changeset_heatmap)

changeset_heatmap_interactive

```


### {.tabset}

```{r}
#| title: "Changesets and size"

changesets_details <- read_csv(paste0(base_path, "data/raw/changesets_details.csv"))

hist_changesets_size <- changesets_details |> 
  count(changeset, name = "n_features") |>
  mutate(changeset = as.numeric(changeset)) |> 
  right_join(changesets, by = join_by(changeset == id)) |> 
  select(changeset, n_features, area_meters) |> 
  mutate(area_hectare = area_meters * 0.0001,
         area_hectare = as.numeric(area_hectare)) |> 
  select(n_features, area_hectare) |> 
  pivot_longer(c(n_features, area_hectare)) |> 
  ggplot(aes(x = value, fill = name)) +
  geom_histogram() +
  facet_wrap(~name, ncol=1, scales = "free_x")+ 
  labs(y = "", x = "") +
  theme_minimal() +
  theme(
    axis.title.y=element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    panel.grid= element_blank()
  ) 

ggplotly(hist_changesets_size)


```

## Row

### What {.tabset}

```{r}
#| title: "Most Used keys"

if (use_db) {
  cat('::: {.callout-warning}\n**Feature keys panel** — available only when `use_db_overlay: false` (API/planet path).\n:::\n')
} else {
  #Non-db-overlay option

  key_counts <- changesets_tags  |>
    count(key, sort = TRUE) |> 
    categorise_keys() |> 
    mutate(parent_key = replace_na(parent_key, "other"),
           top_key = replace_na(top_key, "other"))
  
  top_keys <- levels(as.factor(key_counts$top_key))
  top_keys_n <- nlevels(as.factor(key_counts$top_key))
  
  
  tags_top_parents <- data.frame(
    key = top_keys,
    n = rep(0, top_keys_n),
    top_key = rep("", top_keys_n) 
  )
  
  key_counts <- key_counts |> 
    bind_rows(tags_top_parents) |> 
    filter(key != "leisure")
  
  plot_ly(
    data = key_counts,
    type = "treemap",
    labels = ~key,
    values = ~n,
    parents = ~top_key,
    hoverinfo = "label+value+percent parent+percent entry+percent root"
  )
}

```


```{r}
#| title: "Hashtags used"
#| label: hashtags-count2

# Build a tidy table of hashtag tokens, preferring the overlay when available
if (use_db &&
    exists("changesets_tags") &&
    all(c("changeset","key","value") %in% names(changesets_tags))) {

  ht <- changesets_tags |>
    dplyr::filter(key == "hashtags", !is.na(value), nzchar(value)) |>
    dplyr::transmute(
      hashtag = value |>
        stringr::str_trim() |>
        stringr::str_remove("^#") |>
        tolower()
    )

} else if ("hashtags" %in% names(changesets)) {

  ht <- changesets |>
    dplyr::select(hashtags) |>
    dplyr::filter(!is.na(hashtags), nzchar(hashtags)) |>
    tidyr::separate_rows(hashtags, sep = "[;,]") |>
    dplyr::transmute(
      hashtag = hashtags |>
        stringr::str_trim() |>
        stringr::str_remove("^#") |>
        tolower()
    )

} else {
  ht <- tibble::tibble(hashtag = character(0))
}

if (nrow(ht)) {
  set.seed(42)

  word_corpus <- ht |>
    dplyr::filter(nzchar(hashtag)) |>
    dplyr::count(hashtag, sort = TRUE, name = "n") |>
    dplyr::slice_max(n, n = 50)

  wordcloud::wordcloud(
    words = word_corpus$hashtag,
    freq  = word_corpus$n,
    min.freq = 1,
    random.order = FALSE,
    colors = RColorBrewer::brewer.pal(8, "PuOr"),
    scale = c(2, 0.5)
  )
} else {
  cat("No hashtags found in the input data.")
}
```



```{r}
#| title: "Comments"
library(tidytext)

changesets |>
  select(comment) |> 
  unnest_tokens(bigram, comment, token = "ngrams", n = 2) |> 
  separate(bigram, c("word1", "word2"), sep = " ") |> 
  filter(!word1 %in% stop_words$word,
         !word2 %in% stop_words$word,
         !is.na(word1),
         !is.na(word2)) |> 
  mutate(bigram = paste(word1, word2)) |> 
  count(bigram, sort = TRUE) |> 
  head(15) |> 
  kable()
```

```{r}
#| title: "Top words"

if ("comment" %in% names(changesets)) {
  library(tidytext)

  custom_stop <- c("https","http","amp","osm","map","maps","etc","road","roads",
                   "added","adding","update","updated","updating","fix","fixed",
                   "note","notes")

  words_top <- changesets |>
    dplyr::select(user, created_at, comment) |>
    dplyr::filter(!is.na(comment), nzchar(comment)) |>
    tidytext::unnest_tokens(word, comment, token = "words") |>
    dplyr::mutate(
      word = stringr::str_replace_all(word, "#", ""),
      word = stringr::str_replace_all(word, "[0-9]+", "")
    ) |>
    dplyr::filter(
      !word %in% tidytext::stop_words$word,
      !tolower(word) %in% custom_stop,
      nchar(word) >= 3
    ) |>
    dplyr::count(word, sort = TRUE) |>
    dplyr::slice_head(n = 30)

  if (nrow(words_top)) {
    p <- words_top |>
      dplyr::mutate(word = forcats::fct_reorder(word, n)) |>
      ggplot2::ggplot(ggplot2::aes(x = n, y = word)) +
      ggplot2::geom_col() +
      ggplot2::labs(x = "Count", y = NULL, title = "Most frequent words in comments") +
      ggplot2::theme_minimal()

    plotly::ggplotly(p)
  } else {
    cat("No words to show after filtering.")
  }
} else {
  cat("No 'comment' column found.")
}
```

```{r}
#| title: "Distinctive by user (TF-IDF)"

if ("comment" %in% names(changesets)) {
  library(tidytext)
  library(DT)

  tfidf_tbl <- changesets |>
    dplyr::select(user, comment) |>
    dplyr::filter(!is.na(comment), nzchar(comment), !is.na(user), nzchar(user)) |>
    tidytext::unnest_tokens(word, comment) |>
    dplyr::filter(!word %in% tidytext::stop_words$word, nchar(word) >= 3) |>
    dplyr::count(user, word, sort = TRUE) |>
    tidytext::bind_tf_idf(word, user, n) |>
    dplyr::group_by(user) |>
    dplyr::slice_max(order_by = tf_idf, n = 10, with_ties = FALSE) |>
    dplyr::ungroup() |>
    dplyr::arrange(user, dplyr::desc(tf_idf))

  if (nrow(tfidf_tbl)) {
    DT::datatable(tfidf_tbl, options = list(pageLength = 25), rownames = FALSE)
  } else {
    cat("No distinctive terms to show.")
  }
} else {
  cat("No 'comment' column found.")
}
```


```{r}
#| title: "Comments (raw)"
if ("comment" %in% names(changesets)) {
  library(DT)

  cm_raw <- changesets |>
    dplyr::select(id, user, created_at, comment) |>
    dplyr::filter(!is.na(comment), nzchar(comment)) |>
    dplyr::arrange(dplyr::desc(created_at)) |>
    dplyr::slice_head(n = 20)   # <-- the missing pipe was here

  DT::datatable(
    cm_raw,
    options = list(pageLength = 20, lengthMenu = c(20, 50, 100, 200)),
    rownames = FALSE
  )
} else {
  cat("No 'comment' column found.")
}
```

```{r}
#| title: "Most Used keys 2"


if (use_db) {
  cat('::: {.callout-warning}\n**Feature keys panel** — available only when `use_db_overlay: false` (API/planet path).\n:::\n')
} else {
  #Non-db-overlay option
  key_counts2 <- changesets_tags  |>
    count(key, sort = TRUE) |> 
    categorise_keys() |> 
    count(top_key, parent_key, wt = n) |> 
    mutate(parent_key = replace_na(parent_key, "other"),
           top_key = replace_na(top_key, "other"),
           top_key = fct_infreq(as.factor(top_key))) 
  
  
  tags_treemap <- ggplot(key_counts2, aes(area = n, fill = top_key, 
                          subgroup = top_key,
                          label = parent_key)) +
    geom_treemap(color = "white") +
    geom_treemap_subgroup_border(color = "white") +
    geom_treemap_text(fontface = "italic", colour = "white", alpha = 0.5,
                      place = "centre",
                      grow = TRUE) +
    geom_treemap_subgroup_text(place = "bottom", grow = T, alpha = 1, colour =
                                 "#333333", fontface = "italic", min.size = 0) +
    labs(fill = "") +
    scale_fill_brewer(palette = "Paired", type = "qual", direction = 1) +
    # TODO: sort sorting for scale fill. Now based on factor's levels.
    theme(legend.position="none")
  
  tags_treemap
}

```


### How {.tabset}


```{r}
#| title: "Sources of information"

df <- changesets |>
  dplyr::transmute(
    # force both to character and treat "" as NA
    source = dplyr::na_if(as.character(source), ""),
    imagery_used = dplyr::na_if(as.character(imagery_used), "")
  ) |>
  dplyr::mutate(source = dplyr::coalesce(source, imagery_used)) |>
  dplyr::filter(!is.na(source), nzchar(source)) |>
  # split multi-valued sources like "Bing;Maxar"
  tidyr::separate_rows(source, sep = ";") |>
  dplyr::mutate(
    source = stringr::str_trim(source),
    source = stringr::str_replace_all(source, "\\+", " "),
    source = dplyr::case_when(
      stringr::str_to_lower(source) == "survey" ~ "Survey",
      stringr::str_detect(stringr::str_to_lower(source), "mapillary") ~ "Mapillary",
      stringr::str_detect(stringr::str_to_lower(source), "esri") ~ "Esri World Imagery",
      stringr::str_detect(source, "Bing") ~ "Bing Maps Aerial",
      stringr::str_detect(source, "PNOA") ~ "PNOA Spain",
      stringr::str_detect(source, "Maxar") ~ "Maxar Premium Imagery",
      stringr::str_detect(source, "Catastro") ~ "Catastro Spain",
      stringr::str_detect(source, "Strava") ~ "Strava Heat Map",
      TRUE ~ source
    ),
    parent = dplyr::case_when(
      stringr::str_detect(source, "Bing|Esri|Maxar|PNOA") ~ "Satellite Imagery",
      stringr::str_detect(stringr::str_to_lower(source), "imagery|copernicus") ~ "Satellite Imagery",
      TRUE ~ NA_character_
    )
  ) |>
  dplyr::mutate(source = forcats::fct_lump_n(as.factor(source), 10)) |>
  dplyr::count(source, parent, sort = TRUE)

plotly::plot_ly(df, type = "sunburst",
                labels = ~source, values = ~n, parents = ~parent)


```


```{r}
#| title: "Software used"

changesets |> 
  select(created_by) |> 
  rename(software = created_by) |> 
  mutate(software_clean = case_when(
    str_detect(software, "iD ") ~ "iD",
    str_detect(software, "JOSM") ~ "JOSM",
    str_detect(software, "maps.me") ~ "maps.me",
    str_detect(software, "MapComplete") ~ "MapComplete",
    str_detect(software, "Organic Maps") ~ "Organic Maps",
    str_detect(software, "RapiD") ~ "RapiD",
    str_detect(software, "StreetComplete") ~ "StreetComplete",
    str_detect(software, "Vespucci") ~ "Vespucci",
    .default = "Other"
  )) |> 
  count(software_clean, sort = TRUE) |> 
  mutate(software_clean = fct_reorder(software_clean, n)) |> 
  plot_ly(labels = ~software_clean, values = ~n) |> 
  add_pie(hole = 0.6)


```

```{r}
#| title: "Type of changesets"

# Prefer actions from changesets_details if available; otherwise fall back,
# and if nothing exists, show a friendly placeholder.
if (exists("changesets_details") && "action" %in% names(changesets_details)) {
  df_actions <- changesets_details |>
    dplyr::count(action, sort = TRUE) |>
    dplyr::rename(action_type = action)
} else if ("action_type" %in% names(changesets_tags)) {
  df_actions <- changesets_tags |>
    dplyr::count(action_type, sort = TRUE)
} else {
  df_actions <- tibble::tibble(action_type = "(not available)", n = 1L)
}

df_actions |>
  plotly::plot_ly(labels = ~action_type, values = ~n) |>
  plotly::add_pie(hole = 0.6)

```

```{r}
#| title: Language

get_language_counts <- function(changesets, changesets_tags) {
  # Case 1: locale column present on changesets
  if ("locale" %in% names(changesets) && any(!is.na(changesets$locale))) {
    changesets |>
      dplyr::select(locale) |>
      tidyr::separate_wider_delim(
        locale, delim = "-", names = c("language", "country"),
        too_few = "align_start", cols_remove = FALSE
      ) |>
      dplyr::filter(!is.na(language), nzchar(language)) |>
      dplyr::count(language, sort = TRUE)
  # Case 2: try to recover from tags (locale/language keys)
  } else if (exists("changesets_tags") &&
             all(c("key","value") %in% names(changesets_tags)) &&
             nrow(changesets_tags)) {
    lang_by_tag <- changesets_tags |>
      dplyr::filter(key %in% c("locale","language")) |>
      dplyr::mutate(language = tolower(sub("-.*$", "", value))) |>
      dplyr::filter(!is.na(language), nzchar(language)) |>
      dplyr::count(language, sort = TRUE)
    if (nrow(lang_by_tag)) lang_by_tag else tibble::tibble(language="(not available)", n=1L)
  } else {
    tibble::tibble(language="(not available)", n=1L)
  }
}

lang_counts <- get_language_counts(changesets, changesets_tags)

lang_counts |>
  dplyr::mutate(language = forcats::fct_reorder(language, n)) |>
  plotly::plot_ly(labels = ~language, values = ~n) |>
  plotly::add_pie(hole = 0.6)

```






# Wiki
```{r}
wiki_stats <- calc_stats_contributions_wiki(wiki_contributions) 
```

## Valueboxes

```{r}
#| content: valuebox
#| title: "# users with wiki edits"
#| icon: wikipedia
#| color: primary
wiki_stats[1]
```

```{r}
#| content: valuebox
#| title: "total additions"
#| icon: database-add
#| color: green

df <- as.data.frame(wiki_stats[4])
round(sum(df$additions, na.rm = TRUE), digits = 2)

  
```

```{r}
#| content: valuebox
#| title: "total deletions"
#| icon: database-dash
#| color: red

df <- as.data.frame(wiki_stats[4])
round(sum(df$deletions, na.rm = TRUE), digits = 2)

```

```{r}
#| content: valuebox
#| title: "# of new pages created"
#| icon: file-earmark-plus

# sum(wiki_contributions$new, na.rm = TRUE)
wiki_contributions |> 
  filter(parentid == 0) |> 
  count(parentid) |> 
  pull(n)


```


## Row


```{r}
#| title: When do they contribute?

# Modified - OSM_Aug_RA
# wiki_edits_day_time <- wiki_contributions |> 
#   select(timestamp) |> 
#   mutate(date = ymd_hms(timestamp),
#          weekday = weekdays(date),
#          weekday = forcats::fct_relevel(
#            as.factor(weekday), 
#            "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"),
#          time = hms::as_hms(date),
#          hour = hour(time)) |> 
#   count(weekday, hour) |> 
#   # Converting to wide to generate NAs
#   pivot_wider(names_from = hour, values_from = n) |>
#   pivot_longer(-weekday, names_to = "hour", values_to = "n") |> 
#   mutate(hour = as.numeric(hour)) 
# # mutate(n = replace_na(n, 0))
# 
# heatmap <- ggplot(wiki_edits_day_time, aes(hour, weekday, fill= n)) + 
#   geom_tile(show.legend = FALSE) +
#   labs(y = NULL) + 
#   scale_fill_distiller(palette = "Blues", direction = 1) +
#   # scale_x_continuous(position = "top")  + # Does not work with plotly!
#   theme_minimal() 
# # theme(panel.background = element_rect(fill = 'lightgrey'),
# #       # panel.grid.minor = element_line(color = 'white', size = 2)
# #       )
# 
# heatmap_interactive <- ggplotly(heatmap)
# 
# heatmap_interactive

wiki_edits_day_time <- wiki_contributions |>
  select(timestamp) |>
  mutate(
    date = ymd_hms(timestamp),
    month = factor(month(date), levels = c(1:12), labels = month.abb),
    weekday = weekdays(date),
    weekday = forcats::fct_relevel(
      as.factor(weekday),
      "Monday",
      "Tuesday",
      "Wednesday",
      "Thursday",
      "Friday",
      "Saturday",
      "Sunday"
    )
  ) |>   
  count(month, weekday)


# plotly.js does not (yet) support horizontal legend items, hence legend is
# vertical on the side
wiki_heatmap <-
  ggplot(wiki_edits_day_time, aes(month, weekday, fill = n)) +
  geom_tile() +
  # subtitle not working as expected
  labs(x = NULL, y = NULL) +
  scale_fill_gradient2(
    low = "#ffffff",
    mid = "#d8daeb",
    high = "#2d004b",
    name = "Number of Edits"
  ) +
  scale_y_discrete(limits = rev) +
  #coord_fixed() +
  theme_heatmap

wiki_heatmap_interactive <- ggplotly(wiki_heatmap)

wiki_heatmap_interactive

```

```{r}
#| title: "Written words by language"
wiki_contributions |> 
  select(title, sizediff) |> 
  mutate(lang = str_extract(title, "^([a-zA-Z]){2}\\:"),
         lang = str_remove(lang, ":"),
         lang = str_to_upper(lang),
         lang = case_when(is.na(lang) ~ "EN",
                          .default = lang)) |> 
  count(lang, wt = sizediff, sort = TRUE) |> 
  mutate(lang = fct_reorder(lang, n)) |> 
  plot_ly(labels = ~lang, values = ~n) |> 
  add_pie(hole = 0.6)
```

```{r}
#| title: "Wiki changes and size"
p <- wiki_contributions |> 
  select(size, sizediff) |> 
  # mutate(sizediff = abs(sizediff)) |> 
  ggplot(aes(x = size, y = sizediff)) +
  geom_point() +
  # scale_x_log10() +
  # scale_y_log10() + 
  labs(x = "Page size",
       y = "Characters added/remov") + 
  theme_minimal()

ggplotly(p)
```


## Row
```{r}
#| title: "Top 15 contributed wiki pages"
plot <- wiki_stats[[5]] |>
  mutate(title = fct_reorder(title, n)) |> 
  head(15) |> 
  ggplot(aes(x = n, y = title, label = n)) +
  geom_bar(stat="identity") +
  # geom_text(color = "white", hjust = 1.2) +
  labs(x = NULL, y = NULL) +
  theme_minimal() +
  theme(panel.grid= element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x = element_blank())
 

ggplotly(plot)

```

# Other contributions {orientation="columns"}

## Diaries
```{r}
#| content: valuebox
#| title: "# users with diary entries"
#| icon: pencil-square
#| color: primary
#| height: 30px
contributions_summary |> 
  filter(diary > 0) |> 
  nrow()
```

Total diary entries

Wordcount based on titles
## Notes

```{r}
#| content: valuebox
#| title: "# users with map notes"
#| icon: sticky
#| color: primary
notes_users <- contributions_summary |> 
  filter(map_notes > 0) |> 
  nrow()

notes_users
```

```{r}
#| content: valuebox
#| title: "# of notes"
#| icon: sticky
#| color: primary
notes_total <- contributions_summary |> 
  select(map_notes) |> 
  pivot_longer(map_notes) |> 
  count(name, wt = value) |> 
  pull(n)

format(notes_total, big.mark = ",")
```

```{r}
#| content: valuebox
#| title: "# of notes/user"
#| icon: sticky
#| color: primary
notes_total / notes_users
```

Wordcout based on notes contents

WordLength by notes?


## Traces

```{r}
#| content: valuebox
#| title: "# users with gpx traces"
#| icon: pin-map
#| color: primary
contributions_summary |> 
  filter(traces > 0) |> 
  nrow()
```

### 

Total km

### 
Total traces by country

# About {scrolling="true"}

::: {.card .flow}


This dashboard is a proof of concept that displays how a group of users contribute to OpenStreetMap with special consideration to EDI issues. That means, for example, that this dashboard recognises different types of contributions and all data shown has been anonymised.

This dashboard is a result of a codesign process with members of [Geochicas](https://wiki.openstreetmap.org/wiki/LatAm/Groups/GeoChicas) and the [Centre for Interdisciplinary Methodologies](https://warwick.ac.uk/cim) at the University of Warwick, and is inspired by Pascal Neis' great [HDYC](https://hdyc.neis-one.org/), but with notable important differences: 

- It displays how certain groups of users contribute to OSM, without revealing who their members are. For the purposes of this dashboard, a group is a number of users that have a shared interests who self-report as members of a group.
- It recognises that there are different ways of contributing to OSM beyond map contributions (e.g. wiki edits, user diaries...). As such, it incorporates data from different sources (see below).
- It is opensource: you can see the code (and contribute to it!) in [this GitHub repository](https://github.com/WarwickCIM/OSMdashboard).
- It is self-hosted (for now! -keep tuned), meaning that data stays in your computer.
- It is way slower! 

## Data Sources

- OpenStreetMap API (via [`{{osmapiR}})`](https://docs.ropensci.org/osmapiR/index.html))
- OpenStreetMap Wiki API 
- OpenStreetMap User profiles

## Acknowledgements

## How to cite

If you want to cite this work, please do it as follows:

:::

