---
title: "Feminist Design Principles in OSMdashboard"
bibliography: "references.bib"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(OSMdashboard)
```

This package ins highly inspired and motivated by the 6 design principles for feminist data visualisations proposed by Catherine d'Ignazio and Lauren F. Klein [-@dignazio2023; -@dignazio2016]. In this page we explain how we have implemented (or tried to) each of these principles in the design of this package and its main output, the dashboard.

## Rethinking Binaries

> A feminist approach to data visualization should therefore emphasize representational strategies premised on multiplicity rather than binaries, and acknowledge the limits of any binaristic view [53]. -- @dignazio2016

## Embrace pluralism

> We believe that self-disclosure, and an embrace of pluralism more generally, can do more; it can help to encourage alternatives to the single “view from nowhere” so often favored in visualization design [21]. Ideally, a focus on pluralism would help visualization research move away from its current emphasis on “objective” presentation in favor of designs that facilitate pathways to multiple truths. -- @dignazio2016

We do so by:

- This package is the result of a participatory design process with members of Geochicas and researchers from the University of Warwick, who happen to be long-term OSM contributors.
- Incorporating a diversity of contribution types (see `vignette::dashboard-group-contributions`)
- 

## Examine Power and Aspire to Empowerment

> A feminist approach to data visualization therefore acknowledges the user as a source of knowledge in the design as well as the reception of any visual interface. The creation of knowledge is, after all, always a shared endeavor.

## Consider context

> A feminist approach to data visualization must therefore consider how diverse contexts can influence the production of a visualization, and think through the various ways in which any particular visualization output might be received.

## Legitimize Embodiment and Affect

> With the rise of popular forms of visualization such as data journalism, designers have begun to intentionally leverage affect in order to create an emotional bond with a story or issue [11], or to engage and impress readers with beauty and complexity 61. These affective dimensions of visualization have been under-explored in traditional visualization research. Acknowledging the importance of embodiment and affect also has implications for how we evaluate visualizations.

## Make labor visible

> “Starting with questions of data provenance helps to credit the bodies that make visualization possible – the bodies that collect the data, that digitize them, that clean them, and that maintain them. However, most data provenance research focuses on technical rather than human points of origination and integration [66]. With its emphasis on under-valued forms of labor, a feminist approach to visualization can help to render visible the bodies that shape and care for data at every stage of the process. This relates to the concept of provenance rhetoric [44] in which authors of narrative visualizations cite data sources and methods which may help build credibility with the audience.” -- @dignazio2016, p. 3

- This project welcomes any type of contributions, not just coding. It follows the [all-contributors](https://allcontributors.org) specification as a way to recognise that, while crediting all contributors.
- All contributors are listed in the project's README file. All the authors, maintainers and collaborators of this package are listed in the DESCRIPTION file, and the package's website (https://warwickcim.github.io/OSMdashboard/).

## References
